---
title: "ML_Trial"
author: "Andrea,Gloria,Lorenzo,Thomas"
date: "`r Sys.Date()`"
output: html_document
---
# Preparatory steps
  Loading packages
  Store package names in a vectors for ease of access and to load them easily 
```{r, message=FALSE, warning=FALSE}
# Set the root directory for knitr
knitr::opts_knit$set(root.dir = "D:/VarieTHOM/University/QCB/3_SEMESTRE/Data Mining/Laboratory (Blanzieri)/0_PROJECT/Datasets_finals")

# Define a list of packages to be loaded
PACKAGES <- c(
  "randomForest",    # For random forest modeling
  "caret",           # For machine learning model training and evaluation
  "gmodels",         # Provides tools for model diagnostics
  "yardstick",       # For model evaluation metrics
  "ggplot2",         # For data visualization
  "corrplot",        # For correlation plot visualization
  "tidyverse",       # Collection of packages for data manipulation and visualization
  "factoextra",      # Additional tools for clustering and factor analysis
  "FactoMineR",      # For exploratory data analysis and multivariate analysis
  "glue",            # For string interpolation
  "plotly",          # Interactive plots
  "dplyr",           # To manage data frames more efficiently
  "kernlab",         # Kernel-based machine learning methods
  "readr",           # For reading and parsing data
  "e1071",           # For SVM modeling
  "class",           # Various classification methods
  "MASS"             # Modern Applied Statistics with S
)

# Load the specified packages
invisible(lapply(PACKAGES, library, character.only = TRUE))


```

## Managing the data to use

```{r,warning=FALSE}
# Load the data
setwd("D:/VarieTHOM/University/QCB/3_SEMESTRE/Data Mining/Laboratory (Blanzieri)/0_PROJECT/Datasets_finals")
ML_data <- data.frame(read_csv("ML_data.csv", col_names = TRUE))

# Set rownames
ML_data <- subset(ML_data, select = -((ncol(ML_data) - 4):ncol(ML_data)))
row.names(ML_data) <- ML_data$...1

# Eliminate the columns that are not needed
my_data <- subset(ML_data, select = -c(...1,risk...127))
#my_data$risk...127 <- NULL

# Subset data frame by condition
df_Tumor <- subset(my_data, C_T...125 == "Tumor")
```

## NUMBER COATING

```{r First dumb try }
# Getting ready to do predictions
# Number coating the values
## Specify the columns to be label encoded
columns_to_encode <- c("type...126", "C_T...125", "Cell_type...128")
columns_to_encode2 <- c("type...126", "C_T...125")

# Create a new data frame without rows where 'Cell_type...128' has value "Unknown"
my_data_train <- subset(my_data, Cell_type...128 != "Unkown")

# Convert specified columns to factor type
my_data_train <- my_data_train %>% mutate_at(columns_to_encode, as.factor)

# Display levels of 'Cell_type...128'
levels(my_data_train$Cell_type...128)

# Convert specified columns to factor type for the entire dataset
my_data_encoded <- my_data %>% mutate_at(columns_to_encode, as.factor)


# Do the number coating as numbers
encoded <- my_data_encoded %>% mutate_at(columns_to_encode, as.numeric)
train_data <- my_data_train %>% mutate_at(columns_to_encode, as.numeric)

# Create a dictionary-like structure to store the labels
## The order corresponds to the number
my_levels <- list(
  type...126 = levels(my_data_encoded$type...126),
  C_T...125 = levels(my_data_encoded$C_T...125),
  Cell_type...128 = levels(my_data_encoded$Cell_type...128)
)

my_levels

```

# Split the data sets to start the following analysis

```{r Split the data}
# Use 70% of the dataset as the training set and the remaining 30% as the testing set
# Create a logical vector for sampling with replacement
sample <- sample(c(TRUE, FALSE), nrow(my_data_train), replace = TRUE, prob = c(0.7, 0.3))

# Create the training set
train <- my_data_train[sample, ]
train_lda <- train[, -124]

# Create the testing set
test <- my_data_train[!sample, ]

# Create a logical vector for sampling without replacement for numeric data
sample_numeric <- sample(c(TRUE, FALSE), nrow(train_data), replace = TRUE, prob = c(0.7, 0.3))

# Create the numeric training set
train_n <- train_data[sample, ]

# Create the numeric testing set
test_n <- train_data[!sample, ]

```

# Optimize RF

```{r}
set.seed(1235)
model_tuned <- tuneRF(
               x=my_data_train, #define predictor variables
               y=my_data_train$Cell_type...128, #define response variable
               ntreeTry=800,
               mtryStart=4, 
               stepFactor=1.5,
               improve=0.01,
               trace=FALSE, #don't show real-time progress
               plot = TRUE,
               importance = TRUE
               )
```

```{r Fit the Random Forest model}
# Fit the Random Forest model
set.seed(1234)
model_RF <- randomForest(Cell_type...128 ~ ., 
                      data = train,
                      ntree = 800,
                      importance = TRUE,
                      mtry = 9
)

# Create a copy of the data for predictions
predicted_RF <- my_data_encoded

# Make predictions on the test set
test_RF <- predict(model_RF, newdata = test)

# Fit another Random Forest model on the entire training dataset
model_RF2 <- randomForest(Cell_type...128 ~ ., 
                      data = my_data_train,
                      ntree = 800,
                      importance = TRUE,
                      mtry = 9
)

# Make predictions on the new data
Prediction_RF <- predict(model_RF2, newdata = predicted_RF)

# View predictions
predict_RF <- data.frame(Prediction_RF)
predicted_RF$predicted_rf <- predict_RF$Prediction_RF
setwd("D:/VarieTHOM/University/QCB/3_SEMESTRE/Data Mining/Laboratory (Blanzieri)/0_PROJECT/Datasets_finals")
write.csv(predicted_RF, "Predict_rf.csv", row.names = TRUE)

# Display the predictions
predicted_RF

# Display fitted models
print(model_RF)   # Using print for a more informative display
print(model_RF2)

```

# First basic plots RF

```{r, warning=FALSE}

# Find RMSE of the best model
# sqrt(model$mse[best_trees])  # Uncomment if you need to print the RMSE, works only on fully numbers

# Plot the test MSE by the number of trees
plot(model_RF2, main = "Test MSE by Number of Trees", type = "l", pch = 16)

# Produce a variable importance plot
varImpPlot(model_RF2, main = "Variable Importance Plot", col = "darkgreen")

# Partial dependence plots illustrate the relationship between a specific feature and the predicted outcome while holding other features constant.
# They are useful for understanding the effect of individual features on the model predictions.
partialPlot(model_RF2, train, PC3, col = "darkgreen", )

# Evaluate accuracy
# correct_predictions <- sum(test_RF == test$Cell_type...128)
# accuracy <- correct_predictions / nrow(test)

# Create a confusion matrix
conf_matrix <- table(test_RF, test_n$Cell_type...128)
conf_matrix

# Calculate metrics
accuracy <- sum(diag(conf_matrix)) / sum(conf_matrix)
precision <- diag(conf_matrix) / rowSums(conf_matrix)
recall <- diag(conf_matrix) / colSums(conf_matrix)
f1_score <- 2 * (precision * recall) / (precision + recall)

# Print the metrics
cat("Accuracy:", accuracy, "\n")
cat("Precision:", mean(precision, na.rm = TRUE), "\n")
cat("Recall:", mean(recall, na.rm = TRUE), "\n")
cat("F1 Score:", mean(f1_score, na.rm = TRUE), "\n")

# Create a bar plot for precision, recall, and F1 score
metrics_df <- data.frame(Metric = c("Precision", "Recall", "F1 Score"),
                         Value = c(mean(precision, na.rm = TRUE),
                                   mean(recall, na.rm = TRUE),
                                   mean(f1_score, na.rm = TRUE)))

# Create a bar plot for precision, recall, and F1 score using plotly
plot_ly(metrics_df, x = ~Metric, y = ~Value, type = "bar", color = ~Metric) %>%
  layout(title = "Metrics", xaxis = list(title = "Metric"), yaxis = list(title = "Value"))

```


#  Linear Discriminant Analysis
  A method you can use when you have a set of predictor variables and you’d like to classify a response variable into two or more classes.

```{r}

# Prepare data for LDA prediction
predict_LDA <- my_data_encoded

# Make this example reproducible
set.seed(1234)

# Fit LDA model
lda_model <- lda(Cell_type...128 ~ ., 
                 data = train_lda)

# Remove the 124th column as it is not informative
my_data_LDA <- my_data_train[,-124]
lda_model2 <- lda(Cell_type...128 ~ ., 
                  data = my_data_LDA)

# View model output
## Prior probabilities of group: Represent proportions of each Species in the training set.
## Group means: Display mean values for each predictor variable for each label.
## Coefficients of linear discriminant: Display linear combination of predictor variables used to form the decision rule of the LDA model.
plot(lda_model2)

# Extract LDA scores
lda_scores <- as.data.frame(predict(lda_model2)$x)

# Use LDA model to make predictions on the test data
## Returns a list with three variables: class, posterior probability, and linear discriminants
predicted_LDA <- predict(lda_model2, predict_LDA)

# Create a dataframe with the results
predicted_LDA <- data.frame(predicted_LDA)
# Insert results into the dataframe to compare
predict_LDA$predicted_LDA <- predicted_LDA$class
setwd("D:/VarieTHOM/University/QCB/3_SEMESTRE/Data Mining/Laboratory (Blanzieri)/0_PROJECT/Datasets_finals")
write.csv(predict_LDA, "Predict_lda.csv", row.names = TRUE)

# Predict using the LDA model on the test set
predict_LDA_train <- predict(lda_model, test)



```

# LDA Visulization

```{r warning=FALSE}

# View predicted class for the first six observations in the test set
head(predict_LDA_train$class)

# View posterior probabilities for the first six observations in the test set
head(predict_LDA_train$posterior)

# View linear discriminants for the first six observations in the test set
head(predict_LDA_train$x)

# Create a confusion matrix
conf_matrix <- table(predict_LDA_train$class, test_n$Cell_type...128)
conf_matrix

# Calculate metrics
accuracy <- sum(diag(conf_matrix)) / sum(conf_matrix)
precision <- diag(conf_matrix) / rowSums(conf_matrix)
recall <- diag(conf_matrix) / colSums(conf_matrix)
f1_score <- 2 * (precision * recall) / (precision + recall)

# Print the metrics
cat("Accuracy:", accuracy, "\n")
cat("Precision:", mean(precision, na.rm = TRUE), "\n")
cat("Recall:", mean(recall, na.rm = TRUE), "\n")
cat("F1 Score:", mean(f1_score, na.rm = TRUE), "\n")

# Create a bar plot for precision, recall, and F1 score
metrics_df <- data.frame(Metric = c("Precision", "Recall", "F1 Score"),
                         Value = c(mean(precision, na.rm = TRUE),
                                   mean(recall, na.rm = TRUE),
                                   mean(f1_score, na.rm = TRUE)))

# Create a bar plot for precision, recall, and F1 score using plotly
plot_ly(metrics_df, x = ~Metric, y = ~Value, type = "bar", color = ~Metric) %>%
  layout(title = "Metrics", xaxis = list(title = "Metric"), yaxis = list(title = "Value"))


```


```{r}
# Define data to plot
lda_plot <- cbind(my_data_LDA, predict(lda_model2)$x)

# Create ggplot2 2D scatter plot
ggplot(lda_plot, aes(LD1, LD2, color = Cell_type...128)) +
  geom_point() +
  ggtitle("LDA 2D Scatter Plot")

# Create plotly 2D scatter plot
plot_ly(lda_plot, x = ~LD1, y = ~LD2, color = ~Cell_type...128, type = "scatter", mode = "markers") %>%
  layout(title = "LDA 2D Scatter Plot")

# Create plotly 3D scatter plot
plot_ly(lda_plot, x = ~LD1, y = ~LD2, z = ~LD3, color = ~Cell_type...128, 
        type = "scatter3d", mode = "markers") %>%
  layout(scene = list(aspectmode = "cube")) %>%
  layout(title = "LDA 3D Scatter Plot")

```
 
 
# Lasso Regression
  Lasso regression is a method we can use to fit a regression model when multicollinearity is present in the data.

```{r}

library(glmnet)

# Create a matrix of predictors
X <- as.matrix(train_data[, -which(names(train_data) == "Cell_type...128")])


# Convert the response variable to a numeric vector
y <- as.numeric(train_data$Cell_type...128)

# Perform k-fold cross-validation to find the optimal lambda value
cv_model <- cv.glmnet(X, y, alpha = 1)

# Print the cross-validated results
print(cv_model)

# Find the optimal lambda value that minimizes test MSE
best_lambda <- cv_model$lambda.min
best_lambda

# Produce a plot of test MSE by lambda value
plot(cv_model)
 
```



```{r}
# Find coefficients of the best model
best_model <- glmnet(X, y, alpha = 1, lambda = best_lambda)
best_model1 <- glmnet(as.matrix(train_n[, -which(names(train_n) == "Cell_type...128")]), 
                      as.numeric(train_n$Cell_type...128), 
                      alpha = 1, 
                      lambda = best_lambda)

# Create a copy of the encoded data
predicted_LASSO <- as.matrix(encoded[, -which(names(encoded) == "Cell_type...128")])

#x_test <- as.matrix(test_n[, -which(names(train_n) == "Cell_type...128")])

# Use lasso regression model to predict the response value
matrix_predict1<- round(predict(best_model1, s = best_lambda, newx = as.matrix(test_n[, -which(names(test_n) == "Cell_type...128")])))
matrix_predict <- data.frame(round(predict(best_model, s = best_lambda, newx = predicted_LASSO)))

# Create a confusion matrix
conf_matrix <- table(matrix_predict1, test_n$Cell_type...128)
conf_matrix

# Calculate metrics
accuracy <- sum(diag(conf_matrix)) / sum(conf_matrix)
precision <- diag(conf_matrix) / rowSums(conf_matrix)
recall <- diag(conf_matrix) / colSums(conf_matrix)
f1_score <- 2 * (precision * recall) / (precision + recall)

# Print the metrics
cat("Accuracy:", accuracy, "\n")
cat("Precision:", mean(precision, na.rm = TRUE), "\n")
cat("Recall:", mean(recall, na.rm = TRUE), "\n")
cat("F1 Score:", mean(f1_score, na.rm = TRUE), "\n")

# Add predicted values to the encoded_LS data frame
predicted_LASSO<- data.frame(predicted_LASSO)
predicted_LASSO$Cell_type...128 <- encoded$Cell_type...128
predicted_LASSO$Predict_LASSO <- matrix_predict$s1
predicted_LASSO
setwd("D:/VarieTHOM/University/QCB/3_SEMESTRE/Data Mining/Laboratory (Blanzieri)/0_PROJECT/Datasets_finals")
write.csv(predicted_LASSO, "Predict_lasso.csv", row.names = TRUE)

# Create a bar plot for precision, recall, and F1 score
metrics_df <- data.frame(Metric = c("Precision", "Recall", "F1 Score"),
                         Value = c(mean(precision, na.rm = TRUE),
                                   mean(recall, na.rm = TRUE),
                                   mean(f1_score, na.rm = TRUE)))

# Create a bar plot for precision, recall, and F1 score using plotly
plot_ly(metrics_df, x = ~Metric, y = ~Value, type = "bar", color = ~Metric) %>%
  layout(title = "Metrics", xaxis = list(title = "Metric"), yaxis = list(title = "Value"))

```

# Multiclass SVM

```{r}
# Train the SVM model
svmfit <- svm(train_n, 
                 train_n$Cell_type...128, 
                 kernel = "radial")

# Print the summary of the model
print(svmfit)

# Extract predictors from the prediction data
x_predict <- encoded

# Make predictions
predictions <- round(predict(svm_model, x_predict))

preditions_train <- round(predict(svm_model, test_n))

# Create a confusion matrix
conf_matrix <- table(preditions_train, test_n$Cell_type...128)
conf_matrix
# Calculate accuracy
accuracy <- sum(diag(conf_matrix)) / sum(conf_matrix)

# Calculate metrics
accuracy <- sum(diag(conf_matrix)) / sum(conf_matrix)
precision <- diag(conf_matrix) / rowSums(conf_matrix)
recall <- diag(conf_matrix) / colSums(conf_matrix)
f1_score <- 2 * (precision * recall) / (precision + recall)

# Print the metrics
cat("Accuracy:", accuracy, "\n")
cat("Precision:", mean(precision, na.rm = TRUE), "\n")
cat("Recall:", mean(recall, na.rm = TRUE), "\n")
cat("F1 Score:", mean(f1_score, na.rm = TRUE), "\n")

# Print accuracy
cat("Accuracy:", accuracy, "\n")

# Add the predictions to the prediction data frame
x_predict$predicted_class <- predictions

# View the updated prediction data frame
#print(x_predict)
```

# XGBoost

```{r}

library(xgboost)

# Step 1: Prepare the data
# Ensure that the last column in 'train_n' is the variable you want to predict
# For XGBoost, the response variable should be a factor for classification problems
# Extract predictor variables and response variable
X_train <- as.matrix(train_n[, -which(names(train_n) == "Cell_type...128")])  # Exclude the response variable
y_train <- as.matrix(train_n$Cell_type...128) -1   # Response variable

# Step 2: Train the XGBoost model
xgb_model <- xgboost(data = as.matrix(X_train), 
                     label = as.matrix(y_train), 
                     objective = "multi:softmax", 
                     num_class = 4,
                     nrounds =500)

# Step 3: Make predictions on the new data
X_predict <- test_n[, -ncol(test_n)]  # Exclude the response variable

# Predict the class labels 
predictions_xgb <- data.frame(predict(xgb_model, as.matrix(X_predict)) +1)

predictions_xgb$predicted_xgb<- predictions_xgb$predict.xgb_model..as.matrix.X_predict.....1

predictions_xgb$predict.xgb_model..as.matrix.X_predict.....1<-NULL


# If you want the class probabilities instead of class labels, use:
# predictions <- predict(xgb_model, as.matrix(X_predict), type = "prob")

print(xgb_model)
summary(xgb_model)

# Add the predictions to your prediction data frame
#my_data_predict$predicted_class <- predictions

# View the predicted results
#print(my_data_predict$predicted_class)

```



```{r}
#'predictions_xgb' is a data frame or a tibble containing the predictions
#'test_n' is your test data

# Create a confusion matrix
conf_matrix <- table(predictions_xgb$predicted_xgb, test_n$Cell_type...128)

# Calculate metrics
accuracy <- sum(diag(conf_matrix)) / sum(conf_matrix)
precision <- diag(conf_matrix) / rowSums(conf_matrix)
recall <- diag(conf_matrix) / colSums(conf_matrix)
f1_score <- 2 * (precision * recall) / (precision + recall)

# Print the metrics
cat("Accuracy:", accuracy, "\n")
cat("Precision:", mean(precision, na.rm = TRUE), "\n")
cat("Recall:", mean(recall, na.rm = TRUE), "\n")
cat("F1 Score:", mean(f1_score, na.rm = TRUE), "\n")

# Create a bar plot for precision, recall, and F1 score
metrics_df_xgb <- data.frame(Metric = c("Precision", "Recall", "F1 Score"),
                         Value = c(mean(precision, na.rm = TRUE),
                                   mean(recall, na.rm = TRUE),
                                   mean(f1_score, na.rm = TRUE)))

# Create a bar plot for precision, recall, and F1 score using plotly
plot_ly(metrics_df, x = ~Metric, y = ~Value, type = "bar", color = ~Metric) %>%
  layout(title = "Metrics", xaxis = list(title = "Metric"), yaxis = list(title = "Value"))

```

# Predict with the XGB model on the full training set

```{r}

# Create a matrix of predictors
X <- as.matrix(train_data[, -which(names(train_data) == "Cell_type...128")])

# Convert the response variable to a numeric vector
y <- as.numeric(train_data$Cell_type...128)

Y_minusuno = as.matrix(train_data$Cell_type...128) -1

xgb_model2 <- xgboost(data = as.matrix(X), 
                     label = as.matrix(Y_minusuno), 
                     objective = "multi:softmax", 
                     num_class = 4,
                     nrounds =800)
print(xgb_model2)
summary(xgb_model2)

#  Make predictions on the new data
xgb_predict <- encoded[, -ncol(encoded)]  # Exclude the response variable

# Predict the class labels 
xgb_predictions <- data.frame(predict(xgb_model2, as.matrix(xgb_predict)) +1)

xgb_predictions$predicted_xgb<- xgb_predictions$predict.xgb_model2..as.matrix.xgb_predict.....1

xgb_predictions$predict.xgb_model2..as.matrix.xgb_predict.....1<-NULL

xgb_predict$Cell_type...128 <- encoded$Cell_type...128
xgb_predict$predicted_xgb<-xgb_predictions$predicted_xgb

setwd("D:/VarieTHOM/University/QCB/3_SEMESTRE/Data Mining/Laboratory (Blanzieri)/0_PROJECT/Datasets_finals")
write.csv(xgb_predict, "Predict_xgb.csv", row.names = TRUE)

```

# Naive Bayes Model

```{r Naive Bayes Model, warning=FALSE}
# Step 1: Prepare Data
features_train <- train[, -ncol(train)]
labels_train <- train[, ncol(train)]

features_predict <- test[, -ncol(test)]

Predict_nb <- my_data_encoded[, -ncol(my_data_encoded)]

# Step 2: Train Naive Bayes Model with Tuned Parameter
nb_model <- naiveBayes(features_train, labels_train, laplace = 0.001)

nb_model2 <- naiveBayes(train_data[, -ncol(train_data)], train_data[, ncol(train_data)], laplace = 0.001)

# Step 3: Make Predictions
predictions_nb <- predict(nb_model, features_predict)
predictions_nb2 <- data.frame(predict(nb_model, Predict_nb))

predictions_nb2$predicted_nb <- predictions_nb2$predict.nb_model..Predict_nb.
predictions_nb2$predict.nb_model..Predict_nb. <- NULL
Predict_nb$Cell_type...128 <- my_data_encoded$Cell_type...128
Predict_nb$predicted_nb <- predictions_nb2$predicted_nb
setwd("D:/VarieTHOM/University/QCB/3_SEMESTRE/Data Mining/Laboratory (Blanzieri)/0_PROJECT/Datasets_finals")
write.csv(Predict_nb, "Predict_nb.csv", row.names = TRUE)
```


```{r Naive Bayes Model}
# Create a confusion matrix
conf_matrix <- table(predictions_nb, test_n$Cell_type...128)
conf_matrix
# Calculate metrics
accuracy <- sum(diag(conf_matrix)) / sum(conf_matrix)
precision <- diag(conf_matrix) / rowSums(conf_matrix)
recall <- diag(conf_matrix) / colSums(conf_matrix)
f1_score <- 2 * (precision * recall) / (precision + recall)

# Print the metrics
cat("Accuracy:", accuracy, "\n")
cat("Precision:", mean(precision, na.rm = TRUE), "\n")
cat("Recall:", mean(recall, na.rm = TRUE), "\n")
cat("F1 Score:", mean(f1_score, na.rm = TRUE), "\n")
# Create a bar plot for precision, recall, and F1 score
metrics_df_nb <- data.frame(Metric = c("Precision", "Recall", "F1 Score"),
                         Value = c(mean(precision, na.rm = TRUE),
                                   mean(recall, na.rm = TRUE),
                                   mean(f1_score, na.rm = TRUE)))

metrics_df_nb
# Create a bar plot for precision, recall, and F1 score using plotly
plot_ly(metrics_df, x = ~Metric, y = ~Value, type = "bar", color = ~Metric) %>%
  layout(title = "Metrics", xaxis = list(title = "Metric"), yaxis = list(title = "Value"))
#nb_model

```

